---
title: "data_cleaning"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Load packages
```{r message=FALSE}
library(dplyr)
library(tidyr)
library(readr)
```
Here we will load our data
```{r message=F}
catch_original<-read_csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1")
```
A pipe is basically saying "and then". Here is an example of how you can see what a pipe does.
```{r}
sum(1:10)
1:10 %>% sum()
```

We can use the select command to remove columns from our dataset. 
```{r}
catch_data<- catch_original %>% 
  select(Region:Chum) # or select(-All, -notesRegCode)
```
 Now we are going to use the mutate function because we noticed that the Chinook column was characters, not numbers.
```{r}
catch_clean<-catch_data %>% 
  mutate(Chinook = as.numeric(Chinook))
```
 It says there is a problem. NAs introduced by coercion.
```{r}
#Where is the problem
i<- which(is.na(catch_clean$Chinook)) #401
#What is there?
catch_clean[i,] # There is an NA
catch_data[i,] # We see the original data had an I there. Because it was scanned in so the 1 came out as an I. 
```
 Now let's clean this properly
```{r}
catch_clean<-catch_data %>% 
  mutate(Chinook = ifelse(Chinook == "I", 1,Chinook)) %>% 
  mutate(Chinook = as.numeric(Chinook))
```
Now we want to put this into long form
```{r}
catch_long<-catch_clean %>% 
  pivot_longer(cols = -c(Region, Year), names_to = "species", values_to = "catch") #Put a minus next to the column names you don't want to pivot with
catch_wide<- catch_long %>% 
  pivot_wider(names_from = "species", values_from = "catch")
```
Let's rename one of the columns
```{r}
catch_long<- catch_long %>% 
  rename(catch_thousands = catch)

catch_long <- catch_long %>% 
  mutate(catch = catch_thousands * 1000) %>% 
  select(-catch_thousands)
```
Summarize by mean and counts
```{r}
mean_region<- catch_long %>% 
  group_by(Region) %>% 
 summarise(catch_mean = mean(catch))%>% 
  arrange(desc(catch_mean)) # sort in descending order

n_region<-catch_long %>% 
  group_by(Region) %>% 
  summarise(n=n()) 

```
Look at just the dat from SSE region
```{r}
SSE_catch<-catch_long %>% 
  filter(Region == "SSE")
table(SSE_catch$Region)
```
Let's cover joins now
```{r}
region_defs <- read_csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1") %>% 
    select(code, mgmtArea)

catch_joined <-left_join(catch_long, region_defs, by=c("Region"="code"))# by is mapping between primary and foreign key
```

Separate and unite columns
```{r}
sites_df <- data.frame(site = c("HAW-101",
                                "HAW-103",
                                "OAH-320",
                                "OAH-219",
                                "MAI-039"),
                       stringsAsFactors = FALSE)

sites_df %>% 
  separate(site, c("island", "site_number"), "-")
```
Here is using unite useful for dates
```{r}
dates_df <- data.frame(year = c("1930",
                                "1930",
                                "1930"),
                       month = c("12",
                                "12",
                                "12"),
                       day = c("14",
                               "15",
                               "16"),
                       stringsAsFactors = FALSE)

dates_df<-dates_df %>% 
  unite(date, year, month, day, sep = "-")
```


