---
title: "Using ggplot"
author: "Rebecca Clement"
date: "3/1/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load libraries
```{r message=F}
library(leaflet)
library(dplyr)
library(tidyr)
library(ggplot2)
library(DT)
library(scales) # install.packages("scales") #Has comma function for setting scales for axis
```

```{r}
data_url <- "https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e"

esc <- tryCatch(
    read.csv("data/escapement.csv", stringsAsFactors = FALSE),
    error=function(cond) {
        message(paste("Escapement file does not seem to exist, so get it from the KNB."))
        esc <- read.csv(url(data_url, method = "libcurl"), stringsAsFactors = FALSE)
        return(esc)
    }
)

head(esc)
```
Take dataset and calculate annual escapement by region and species
So we first need to separate the sampleDate into year
Then we need to group by region and year, and then species and year
```{r message=F}
annual_esc <- esc %>% 
  separate(sampleDate, c("Year", "Month", "Day"), sep = "-") %>% 
  mutate(Year = as.numeric(Year)) %>% 
  group_by(Species, SASAP.Region, Year) %>% 
  summarize(escapement = sum(DailyCount)) %>% 
  filter(Species %in% c("Chinook", "Sockeye", "Chum", "Coho", "Pink"))
```
Make a ggplot
```{r}
ggplot(annual_esc, aes(x=Species, y=escapement, fill=SASAP.Region)) +
  geom_col()
```
Look at just the dataset with Kodiak
```{r}
kodiak_esc<- annual_esc %>% 
  filter(SASAP.Region=="Kodiak")
mytheme<-theme_classic() +
  theme(legend.position = "top",legend.title=element_blank())
ggplot(kodiak_esc, aes(x=Year, y=escapement, color=Species)) +
  geom_line() +
  #geom_point() +
  ylab("Escapement") +
  scale_y_continuous(labels=comma) +
  ggtitle("Kodiak Salmon Escapement") +
  mytheme

```
Using facet wrap
```{r}

my_plot<-ggplot(annual_esc, aes(x=Year, y=escapement, color=Species)) +
  geom_line() +
  geom_point() +
  ylab("Escapement") +
  scale_y_continuous(labels=comma) +
  facet_wrap( ~ SASAP.Region, scales="free_y",ncol=2)+
  mytheme
ggsave("chart.png",my_plot)
```
Show some more dplyr functions to make a nice interactive table
```{r}

locations <- esc %>% 
  distinct(Location, Latitude, Longitude) %>% 
  drop_na() 
datatable(locations)
```
Show an interactive map using leaflet
```{r}
leaflet(locations) %>% 
  addTiles() %>% 
  addMarkers(lng= ~ Longitude, lat = ~ Latitude, popup= ~ Location)
```

Add a background to it
```{r}
leaflet(locations) %>% 
  addWMSTiles("https://www.gebco.net/data_and_products/gebco_web_services/web_map_service/mapserv?",
              layers = 'GEBCO_LATEST',
              attribution = "Imagery reproduced from the GEBCO_2014 Grid, version 20150318, www.gebco.net") %>%
  addCircleMarkers(lng = ~Longitude,
                   lat = ~Latitude,
                   popup = ~ Location,
                   radius = 5,
                   # set fill properties
                   fillColor = "salmon",
                   fillOpacity = 1,
                   # set stroke properties
                   stroke = T,
                   weight = 0.5,
                   color = "white",
                   opacity = 1)
```

