---
title: "Shapefiles"
author: "Rebecca Clement"
date: "3/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load libraries
```{r message=False}
library(sf)
library(dplyr)
library(ggplot2)
library(leaflet)
library(scales)
library(ggmap)
```
Read in shapefile using sf and look at part of it
```{r}
ak_regions <- read_sf("shapefiles/ak_regions_simp.shp")
plot(ak_regions)  
```
Look at data and transform from one coordinate reference system to another. Plot it
```{r}
class(ak_regions)
head(ak_regions)
st_crs(ak_regions) #View what coordinate reference system

ak_regions_3338 <- ak_regions %>%
  st_transform(crs = 3338)

st_crs(ak_regions_3338)
plot(ak_regions_3338)
```

sf works with dplyr!
```{r}
ak_regions_3338 %>%
  select(region)
```
Using filter
```{r}
ak_regions_3338 %>%
  filter(region == "Southeast")
```
Read in population table
```{r}
pop <- read.csv("shapefiles/alaska_population.csv")
```
Turn this into sf object
```{r}
pop_4326 <- st_as_sf(pop, 
                  coords = c('lng', 'lat'),
                  crs = 4326,
                  remove = F)

head(pop_4326)
```

What region does each city live in?
```{r}
pop_joined <- st_join(pop_4326, ak_regions_3338, join = st_within) # error! because crs are not the same.
pop_3338 <- st_transform(pop_4326, crs = 3338)
pop_joined <- st_join(pop_3338, ak_regions_3338, join = st_within)

head(pop_joined)
plot(pop_joined)
```
Group and summarize
```{r}
pop_region <- pop_joined %>% 
  st_drop_geometry() %>% 
  group_by(region) %>% 
  summarise(total_pop = sum(population))

head(pop_region)
```
Get population onto regions file
```{r}
pop_3338<-left_join(ak_regions_3338,pop_region)
plot(pop_3338)
```
Plot population by management area
```{r}
pop_mgmt_3338<-pop_3338 %>% 
  group_by(mgmt_area) %>% 
  summarise(total_pop = sum(total_pop), do_union=FALSE)

plot(pop_mgmt_3338)
```
Making our map
For maps, put aes in geom calls.
```{r}
ggplot() +
  geom_sf(data=pop_3338,aes(fill=total_pop))+
  theme_bw()+
  labs(fill="Total Population")+
  scale_fill_continuous(low="coral2", high="deepskyblue",labels=comma) #The labels comma takes it out of scientific notation
```
Read in rivers shapefile
```{r}
rivers_3338<-read_sf("shapefiles/ak_rivers_simp.shp")
st_crs(rivers_3338)
```
Add rivers to plot
```{r}
ggplot() +
  geom_sf(data=pop_3338,aes(fill=total_pop))+
  geom_sf(data=rivers_3338, aes(size=StrOrder),color="black") + # this looks terrible if you don't filter down to just stream order
  scale_size(range=c(0.01,0.2), guide=F)+
  theme_bw()+
  labs(fill="Total Population")+
  scale_fill_continuous(low="coral2", high="deepskyblue",labels=comma)
```
## Using ggmap
Pseudo-mercator: 3857
```{r}
pop_3857<-pop_3338 %>% 
  st_transform((crs=3857))
```
Make a function that will transform bounding box to compatible with stamen map
```{r}
# Define a function to fix the bbox to be in EPSG:3857
# See https://github.com/dkahle/ggmap/issues/160#issuecomment-397055208
ggmap_bbox_to_3857 <- function(map) {
  if (!inherits(map, "ggmap")) stop("map must be a ggmap object")
  # Extract the bounding box (in lat/lon) from the ggmap to a numeric vector, 
  # and set the names to what sf::st_bbox expects:
  map_bbox <- setNames(unlist(attr(map, "bb")), 
                       c("ymin", "xmin", "ymax", "xmax"))
  
  # Coonvert the bbox to an sf polygon, transform it to 3857, 
  # and convert back to a bbox (convoluted, but it works)
  bbox_3857 <- st_bbox(st_transform(st_as_sfc(st_bbox(map_bbox, crs = 4326)), 3857))
  
  # Overwrite the bbox of the ggmap object with the transformed coordinates 
  attr(map, "bb")$ll.lat <- bbox_3857["ymin"]
  attr(map, "bb")$ll.lon <- bbox_3857["xmin"]
  attr(map, "bb")$ur.lat <- bbox_3857["ymax"]
  attr(map, "bb")$ur.lon <- bbox_3857["xmax"]
  map
}
```
set bounding box
```{r}
bbox<-c(-170,52,-130,64)
ak_map<-get_stamenmap(bbox,zoom=4)
ak_map_3857<- ggmap_bbox_to_3857(ak_map)
```
```{r}
ggmap(ak_map_3857)+
  geom_sf(data=pop_3857,aes(color=population),inherit.aes=FALSE)+
  scale_color_continuous(low="green",high="black", labels=comma)
```

