---
title: "Intro to rmarkdown"
output: html_document
bibliography: references.bib
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readr)
```

## Headers

#### Header at a deeper level

Now I am writing some text to try to figure out what is going on here.

-   This is a list

-   The second thing on list

-   this

    -   list with pluses
    -   second **thing** in list with pluses

*Italic text*

Numbered list

1.  This is first
2.  this is next
3.  this is third

Oh this is very cool, you can say that there are `r nrow(cars)` cars.

I also didn't realize you can end a line\
with two or more spaces

I know how to [make links](google.com)

> What are block quotes I want to know

$equation$

$$ equation $$ Three or more asterisks is a page break

------------------------------------------------------------------------

superscript^2^

subscript~2~

~~strikethrough~~

You can use the funky visual method to do citations. [@Yashiro2019]

| Species          | Number | Drawing |
|------------------|--------|---------|
| Anax strenuus    | 48     | no      |
| Anax walsinghami | 15     |         |
|                  |        |         |
|                  |        |         |
|                  |        |         |

: This is a table of things we do



#### First code chunk
Command-option-i will insert a new code chunk for you
```{r}
x<-mean(c(1,3,2,4,7))
print(x)

```


```{r}
carsy<-cars[4,]
head(carsy)
```

```{r}
print("I can do this!")
```

## Introduction  
* read in and reference data source  
```{r Read in files, message=FALSE}
bg_chem<-read_csv("BGchem2008data.csv")
head(bg_chem)
```

## Analysis  
* Calculate basic summary stats
```{r}
NO3_mean<-mean(bg_chem$NO3)
NO2_mean<-mean(bg_chem$NO2)
NH4_mean<-mean(bg_chem$NH4)
P_mean<-mean(bg_chem$P)

```
The mean of Nitrate is `r round(NO3_mean,2)`.  
The mean of Nitrite is `r round(NO2_mean,2)`.  
The mean of ammonia is `r round(NH4_mean,2)`.  
The mean of phosphorus is `r round(P_mean,2)`.  

* calculate mean redfield ratio
```{r}
red<-(NO3_mean+NO2_mean+NH4_mean)/P_mean
```
The Redfield ratio is `r round(red,2)`.  

* plot redfield ratio
```{r, echo=F}

plot(bg_chem$P, bg_chem$NO2 + bg_chem$NO3 + bg_chem$NH4)
```

